Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-08-01 analysis/paper/hotspots-paper.Rmd
adding results text

@@

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```

@@

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```

@@

>>>>

The Chr 5 scatter plot of mediation LOD difference values against pleiotropy test statistics 
reveals a block-diagonal pattern of blue coloring among the nonlocal traits. We observe the 
set of nonlocal traits clustered into two blocks. Within each block, any two traits share a 
QTL, ie, are pleiotropic. Yet, the off-diagonal cells, ie, those nonlocal trait pairs that 
include one trait from each of the two blocks, tend to have greater pleiotropy test 
statistics. Still, not all off-diagonal blocks have large pleiotropy test statistics. The 
mostly dark blue matrix on the diagonal (from column 22 to 99) features no cells with pleiotropy test statistics above 3.2. While we didn't determine p-values for these analyses, prior analyses suggest that pleiotropy test statistics below 3.2 are consistent with p-values above 0.05, and, thus, are consistent with a single pleiotropic QTL.

Columns 100 to 119 (and rows 100 to 119) are a second (symmetric) block with only small 
pleiotropy test statistics. Columns 120 to 202 contains cells for the third (symmetric) block. Column 203, the last column, contains a trait *Gm* that, based on pleiotropy test statistics, doesn't cluster near other traits.

@@

@@

Our analysis of the Chromosome 13 hotspot presents results much like those for the Chromosome 2 
hotspot, with the Chromosome 13 local transcript *Il6st* playing a role like that of *Hnf4a* in the 
Chr 2 hotspot. In examining the scatter plot (Figure ) of mediation LOD differences against 
pleiotropy test statistics for all local-nonlocal pairs, we see, as in Figure for the Chr 2 
hotspot, a single local trait *Il6st* has strong mediation effects on a collection of nonlocal traits. *Il6st*, first identified by @keller2018genetic, accounts for most of the nonlocal traits that are affected by the hotspot.

The heatmap for pleiotropy statistics for pairwise tests involving nonlocal traits resembles the 
heatmap for the Chr2 hotspot. Nearly all nonlocal traits have small test statistics when paired 
with each other. This is visualized as a dark blue, symmetric square along the heatmap's diagonal 
with only a few outlier nonlocal traits that demonstrate distinct patterns in pleiotropy test 
statistics.

@@

- lod difference threshold for declaration of "mediation". 
  - esp with regards to Chr 5 hotspot, where the keller mediator has a lot of traits with lod differences under 3, but above the 1.5 threshold.

- look at outputs of `lod_int()`, ie, the support intervals for all traits in a hotspot. Could plot interval against cluster order, for example.
